---
title: "Untitled"
output: html_document
date: "2026-01-18"
---



1. Chi-Square Test of Independence

Use this for: Categorical vs. Categorical relationships.

Application: Test if Gender or Faculty influences Post-Graduation Plans.

Hypothesis Example: H 
0
​	
 : There is no association between a student's faculty and their primary plan after graduation.

Visual Support: This supports any grouped bar charts showing plan distribution by faculty.

2. Kruskal-Wallis Test

Use this for: Ordinal Rankings (like your Question 29).

Application: Test if different Age Groups or Years of Study have significantly different rankings for Work-Life Balance or Salary.

Why this test? Since your rankings are ordinal (1–5) and not normally distributed, this non-parametric alternative to ANOVA is the statistically sound choice for STAT365.

3. One-Way ANOVA (or Kruskal-Wallis for Salary)

Use this for: Categorical vs. Numerical (Salary Expectations).

Application: Test if the Mean Expected Salary differs significantly across Faculties.

Claim Support: This statistically validates your visualization showing if Engineering students have higher salary expectations than other faculties.

4. Correlation Analysis (Spearman's Rho)

Use this for: Numerical vs. Numerical (or Ordinal).

Application: Examine the relationship between Cumulative GPA and Salary Expectation.

Claim Support: It answers whether students with higher academic performance also hold higher financial expectations.
```{r}

# Convert salary to an ordered factor for testing
clean_data$salary_factor <- factor(clean_data$salary_clean, 
                                   levels = c("< 30,000", "30,000–49,999", "50,000–69,999", 
                                              "70,000–99,999", "100,000+"), 
                                   ordered = TRUE)

# Run Kruskal-Wallis Test
k_test <- kruskal.test(as.numeric(salary_factor) ~ faculty_merged, data = clean_data)
print(k_test)

# If significant (p < 0.05), run a post-hoc test to see WHICH faculties differ
if(k_test$p.value < 0.05){
  library(FSA)
  dunnTest(as.numeric(salary_factor) ~ faculty_merged, data = clean_data, method="bh")
}
```




```{r}
# Convert GPA ranges to numeric ranks (1-5)
clean_data$gpa_rank <- as.numeric(factor(clean_data$`6- Please indicate your current cumulative GPA (on a 4.00 scale).`, 
                   levels = c("0–2.0", "2.01–2.5", "2.51–3.0", "3.01–3.5", "3.51–4.0", "Not Applicable")))

# Run Spearman Correlation (Rank-based)
cor_test <- cor.test(clean_data$gpa_rank, as.numeric(clean_data$salary_factor), 
                     method = "spearman", exact = FALSE)
print(cor_test)
```


faculty- career priotries +

```{r}
# Convert the priority ranking to an ordered factor for one variable, e.g., Work-Life Balance
clean_data$WLB_Rank <- factor(clean_data$`29-  Please order the following from 1 (most important) to 5 (least important) according to your priorities. Each number can be used only once. [Work–life balance]`, 
                              levels = c("1", "2", "3", "4", "5"), 
                              ordered = TRUE)

# Run the Kruskal-Wallis Test across Faculty groups
k_test_faculty <- kruskal.test(as.numeric(WLB_Rank) ~ faculty_merged, data = clean_data)
print(k_test_faculty)




# Interpretation: 
# If p-value < 0.05, faculty significantly influences the priority given to Work-Life Balance.

clean_data$salary_benefit <- factor(clean_data$`29-  Please order the following from 1 (most important) to 5 (least important) according to your priorities. Each number can be used only once. [ High salary & benefits]`, 
                              levels = c("1", "2", "3", "4", "5"), 
                              ordered = TRUE)

# Run the Kruskal-Wallis Test across Faculty groups
k_test_faculty <- kruskal.test(as.numeric(salary_benefit) ~ faculty_merged, data = clean_data)
print(k_test_faculty)





clean_data$meaningful_work <- factor(clean_data$`29-  Please order the following from 1 (most important) to 5 (least important) according to your priorities. Each number can be used only once. [ Meaningful work (positive social impact)]`, 
                              levels = c("1", "2", "3", "4", "5"), 
                              ordered = TRUE)

# Run the Kruskal-Wallis Test across Faculty groups
k_test_faculty <- kruskal.test(as.numeric(meaningful_work) ~ faculty_merged, data = clean_data)
print(k_test_faculty)





clean_data$stability <- factor(clean_data$`29-  Please order the following from 1 (most important) to 5 (least important) according to your priorities. Each number can be used only once. [ Career stability]`, 
                              levels = c("1", "2", "3", "4", "5"), 
                              ordered = TRUE)

# Run the Kruskal-Wallis Test across Faculty groups
k_test_faculty <- kruskal.test(as.numeric(stability) ~ faculty_merged, data = clean_data)
print(k_test_faculty)









clean_data$indep <- factor(clean_data$`29-  Please order the following from 1 (most important) to 5 (least important) according to your priorities. Each number can be used only once. [ Autonomy & independence]`, 
                              levels = c("1", "2", "3", "4", "5"), 
                              ordered = TRUE)

# Run the Kruskal-Wallis Test across Faculty groups
k_test_faculty <- kruskal.test(as.numeric(indep) ~ faculty_merged, data = clean_data)
print(k_test_faculty)
```






```{r}
# 1. Prepare the Salary variable as numeric ranks
clean_data$salary_numeric <- as.numeric(factor(clean_data$salary_clean, 
                                        levels = c("< 30,000", "30,000–49,999", "50,000–69,999", 
                                                   "70,000–99,999", "100,000+","Prefer not to say"), 
                                        ordered = TRUE))

# 2. Run One-Way ANOVA for Faculty
anova_faculty <- aov(salary_numeric ~ faculty_merged, data = clean_data)
summary(anova_faculty)

# 3. Run One-Way ANOVA for Post-Graduation Plan
anova_plan <- aov(salary_numeric ~ `13- After completing your degree, what is your primary plan?`, 
                  data = clean_data)
summary(anova_plan)
```



```{r}
# 1. Prepare numeric Salary Rank (1 to 5)
clean_data$salary_num <- as.numeric(factor(clean_data$salary_clean, 
                                    levels = c("< 30,000", "30,000–49,999", "50,000–69,999", 
                                               "70,000–99,999", "100,000+")))

# 2. Prepare numeric GPA Rank (1 to 5)
clean_data$gpa_num <- as.numeric(factor(clean_data$`6- Please indicate your current cumulative GPA (on a 4.00 scale).`,
                                 levels = c("0–2.0", "2.01–2.5", "2.51–3.0", "3.01–3.5", "3.51–4.0")))

# 3. Ensure Gender and Faculty are factors
clean_data$gender_f <- as.factor(clean_data$`5- Which of the following best describes your gender identity?`)
clean_data$faculty_f <- as.factor(clean_data$faculty_merged)

# 4. Run the Multiple Linear Regression
model <- lm(salary_num ~ gpa_num + gender_f + faculty_f, data = clean_data)

# 5. Check results
summary(model)



# Visual check (The 4 standard plots)
par(mfrow = c(2, 2))
plot(model)

```










```{r}
library(car)
# Levene's Test for Homogeneity of Variance
leveneTest(salary_numeric ~ faculty_merged, data = clean_data)
```




```{r}
# Load libraries
library(broom)
library(dplyr)

# Function to test each question against "Neutral" (3)
test_results <- likert_numeric %>%
  pivot_longer(cols = everything(), names_to = "Question", values_to = "Score") %>%
  group_by(Question) %>%
  summarise(
    Mean = mean(Score, na.rm = TRUE),
    SD = sd(Score, na.rm = TRUE),
    # Run the T-Test: Is the mean different from 3?
    p_value = t.test(Score, mu = 3)$p.value,
    # Confidence Interval to show the range
    Conf_Low = t.test(Score, mu = 3)$conf.int[1],
    Conf_High = t.test(Score, mu = 3)$conf.int[2]
  ) %>%
  mutate(Significance = ifelse(p_value < 0.05, "Significant", "Not Significant"))

print(test_results)
```


